Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-14 paper/paper.tex
centering the code

\begin{lstlisting}
    e $\in$ Exp ∂::= ae | (let ([x (ae ae)]) e)
   ae $\in$ AExp ::= ∂x | lam
  lam $\in$ Lam ∂::= (lambda (x) e)
    x $\in$ Var   (variable names)
\end{lstlisting}

@@

\begin{lstlisting}
case class Store[K,V](map: Map[K, Set[V]]) {
  def apply(addr: K): List[V] = map(addr).toList
  def update(addr: K, d: Set[V]): Store[K,V] = {
    val oldd = map.getOrElse(addr, Set())
    Store[K, V](map ++ Map(addr -> (d ++ oldd)))
  }
  def update(addr: K, sd: V): Store[K,V] = update(addr, Set(sd))
}
type BStore = Store[BAddr, Storable]
type KStore = Store[KAddr, Cont]
\end{lstlisting}

@@

\begin{lstlisting}
case class Frame(x: String, e: Expr, env: Env)
case class Cont(frame: Frame, kaddr: KAddr)
\end{lstlisting}

@@

\begin{lstlisting}
def allocBind(x: String, time: Time): BAddr = BAddr(x, time)
def allocKont(tgtExpr: Expr): KAddr = ContAddr(tgtExpr)
\end{lstlisting}

@@

\begin{lstlisting}
def drive(todo: List[State], seen: Set[State]): Set[State] = todo match {
  case Nil => seen
  case hd::tl if seen.contains(hd) => drive(tl, seen)
  case hd::tl => drive(step(hd).toList ++ tl, seen + hd)
}

def analyze(e: Expr): Set[State] = drive(List(inject(e)), Set())
\end{lstlisting}

@@

\textbf{Cache-Passing Style.}
We can now transform our interpreter into cache-passing style. We begin by changing the 
return type @Ans@ to a case class which contains a @VS@ set and a cache. @Ans@ also supports
its own accumulating operation @++@ which defines as union on @Set[VS]@ and join on caches.

\begin{lstlisting}
case class Ans(vss: Set[VS], cache: Cache) {
  def ++(ans: Ans): Ans = { Ans(vss ++ ans.vss, cache.join(ans.cache)) }
}
\end{lstlisting}

>>>>

%------------------------------------------------------------------------------------------------------------
\begin{lstlisting}
                                    ae $\in$ AExp ::= ∂x | lam
                                     x $\in$ Var   (variable names)
                                   lam $\in$ Lam ∂::= (lambda (x) e)
                                     e $\in$ Exp ∂::= ae | (let ([x (ae ae)]) e)
\end{lstlisting}

@@

\begin{lstlisting}
                          case class Store[K,V](map: Map[K, Set[V]]) {
                            def apply(addr: K): List[V] = map(addr).toList
                            def update(addr: K, d: Set[V]): Store[K,V] = {
                              val oldd = map.getOrElse(addr, Set())
                              Store[K, V](map ++ Map(addr -> (d ++ oldd)))
                            }
                            def update(addr: K, sd: V): Store[K,V] = update(addr, Set(sd))
                          }
               type BStore = Store[BAddr, Storable];            type KStore = Store[KAddr, Cont]
\end{lstlisting}

@@

\begin{lstlisting}
     case class Frame(x: String, e: Expr, env: Env);      case class Cont(frame: Frame, kaddr: KAddr)
\end{lstlisting}

@@

\begin{lstlisting}
                           def allocKont(tgtExpr: Expr): KAddr = ContAddr(tgtExpr)
                           def allocBind(x: String, time: Time): BAddr = BAddr(x, time)
\end{lstlisting}

@@

\begin{lstlisting}
              def drive(todo: List[State], seen: Set[State]): Set[State] = todo match {
                case Nil => seen
                case hd::tl if seen.contains(hd) => drive(tl, seen)
                case hd::tl => drive(step(hd).toList ++ tl, seen + hd)
              }
              def analyze(e: Expr): Set[State] = drive(List(inject(e)), Set())
\end{lstlisting}

@@

\textbf{Cache-Passing Style.}
We can now transform our interpreter into cache-passing style. We begin by changing the 
return type @Ans@ to a case class which contains a @VS@ set and a cache. @Ans@ also supports
its own accumulating operation @++@ which defines as union on @Set[VS]@ and join on caches.
% TODO join?
\begin{lstlisting}
                case class Ans(vss: Set[VS], cache: Cache) {
                  def ++(ans: Ans): Ans = { Ans(vss ++ ans.vss, cache.join(ans.cache)) }
                }
\end{lstlisting}
